---
title: "TipBurn WGCNA"
output: html_document
---

```{r}
library(WGCNA)
setwd("/Users/rongkui/Desktop/Lab/TipBurn/WGCNA/")
options(stringsAsFactors = FALSE)
femData = read.csv("all_counts.csv")
dim(femData)
head(femData)
datExpr0 = as.data.frame(t(femData[,-c(1,2)]))
names(datExpr0) = femData$Genes
rownames(datExpr0) = names(femData)[-c(1,2)]
dim(datExpr0)
datExpr0[1:30,1:10]
datExpr0 = datExpr0[-c(8,16),] #taking out B2D28 and B3F30
dim(datExpr0)
datExpr0[1:28,1:10]
#each row is a sample and each column is a gene i think maybe. Anyhow this is a WIDE dataframe. 
```

check if some genes have too many missing datapoint.  
```{r}
gsg = goodSamplesGenes(datExpr0, verbose = 3)
gsg$allOK #you want this to be true

if (!gsg$allOK)
{
# Optionally, print the gene and sample names that were removed:
if (sum(!gsg$goodGenes)>0)
printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
if (sum(!gsg$goodSamples)>0)
printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
# Remove the offending genes and samples from the data:
datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}

dim(datExpr0)  # 28 37734, still a bit too big

#alternatively you could set your own filter for missing data. I feel like I should just because there are way too many genes.   

datExpr0 = datExpr0[,colSums(datExpr0 > 10) >= 7]
dim(datExpr0) #28 23568
```

Next we cluster all the SAMPLES (as oppose to clustering genes that will come later) to see if there are any obvious outliers.  
```{r}
sampleTree = hclust(dist(datExpr0), method = "average")
sizeGrWindow(12,9)
pdf(file = "Fig1.Plots_sampleClustering.pdf", width = 12, height = 9)
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub = "", xlab = "", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
#not sure how reliable this outlier detecter works, cuz it did not detect B2D28 and B3F30 the first time around I had to remove those manually. 
```

Not sure how reliable this outlier detecter works, cuz it did not detect B2D28 and B3F30 the first time around I had to remove those manually. 

```{r}
#abline(h=15, col = "red")
#clust = cutreeStatic(sampleTree, cutHeight = 15, minSize = 10)
#table(clust) #clust 1 contains the samples we want to keep
#keepSamples = (clust==1)
datExpr = datExpr0
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
```

Loading clinical trait data  
```{r}
traitData = read.csv("metadata.csv")
dim(traitData)
names(traitData)
#we can only keep the quantitative ones
traitData = traitData[,-c(2,3,4)]
traitData
#form a data frame analogous to expression data that will hold the clinical traits
Samples = rownames(datExpr)
traitRows = match(Samples, traitData$RNAseq.Sample.Name)
datTraits = traitData[traitRows, -1]
rownames(datTraits) = traitData[traitRows,1]

datTraits

collectGarbage()
```


```{r}
#re-cluster samples
sampleTree2 = hclust(dist(datExpr), method = "average")
#convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(datTraits, signed = FALSE)
plotDendroAndColors(sampleTree2, traitColors, groupLabels = names(datTraits), main = "Sample dendrograms and trait heatmap")
save(datExpr, datTraits, file = "TipBurn-01-dataInput.RData")
```

Yay that was the first session already! Now let's move on to session two: netwrok construction and module detection

```{r}
# choose a set of soft-threshold powers
powers = c(c(1:10), seq(from = 12, to = 40, by = 2))
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
sizeGrWindow(9,5)
par(mfrow = c(1,2))
cex1 = 0.9
#Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2", type="n", main = paste("Scale independence"), col = "black") ;
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers,cex=cex1,col="red") 
abline(h=0.50,col="red")
#make sure to run these three steps together....
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

step-by-step network construction
```{r}
softPower = 20
adjacency = adjacency(datExpr, power = softPower)
TOM = TOMsimilarity(adjacency)
dissTOM = 1-TOM
geneTree = hclust(as.dist(dissTOM), method = "average")
# plot
sizeGrWindow(12, 9)
plot(geneTree, xlab = "", sub = "", main = "Gene clustering on a TOM-based dissimilarity", labels = FALSE, hang = 0.04)

#we like large modules, so we set minimum module size relatively large:
minModuleSize = 50
#Module identification using dynamic tree cut
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize)
table(dynamicMods)
```



```{r}
# Convert labels to colors for plotting
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and the module colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree,dynamicColors, "Dynamic Tree Cut", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, main = "Gene dendrogram and module colors")
```

Merging modules whose expression profiles are very similar:  

```{r}
#calculate eigengenes:
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
#Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs)
METree = hclust(as.dist(MEDiss), method = "average")
sizeGrWindow(7,6)
plot(METree, main = "Clustering of module eigengenes", xlab = "", ylab = "")
MEDissThres1 = 0.08
MEDissThres2 = 0.05
MEDissThres3 = 0.04
MEDissThres4 = 0.03
abline(h = MEDissThres1, col = "red")
abline(h = MEDissThres2, col = "red")
abline(h = MEDissThres3, col = "red")
abline(h = MEDissThres4, col = "red")
merge1 = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres1, verbose = 3)
mergedColors1 = merge1$colors
mergedMEs1 = merge1$newMEs
merge2 = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres2, verbose = 3)
mergedColors2 = merge2$colors
mergedMEs2 = merge2$newMEs
merge3 = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres3, verbose = 3)
mergedColors3 = merge3$colors
mergedMEs3 = merge3$newMEs
merge4 = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres4, verbose = 3)
mergedColors4 = merge4$colors
mergedMEs4 = merge4$newMEs
```

More plotting before and after merging:  
```{r}
sizeGrWindow(12,9)
pdf(file = "Plots_geneDentro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors4, mergedColors3, mergedColors2, mergedColors1), c("Dynamic Tree Cut", "Merged Dynamic 0.03", "Merged Dynamic 0.04", "Merged Dynamic 0.05", "Merged Dynamic 0.08"), dendroLabels = FALSE, hang = 0.03, addGuide = FALSE, guideHang = 0.05)
```

So I think we should just keep things unmerged cuz we lose a lot of information even merging at very low height.  11:30am
OK on a second though (after seeing the module-phenotype correlation matrix) I decided to merge at 0.04.


```{r}
moduleColors = mergedColors3
colorOrder = c("grey", standardColors(50))
moduleLabels = match(moduleColors, colorOrder)-1
MEs = mergedMEs3
save(MEs, moduleLabels, moduleColors, geneTree, file = "TipBurn-02-networkConstruction-SbS.RData")
```


Yay that was part two! Now onto part 3, relating modules to external information and identifying important genes  

```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
#MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
#MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
```

```{r}
sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(", signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor, xLabels = names(datTraits), yLabels = names(MEs), ySymbols = names(MEs), colorLabels = FALSE, colors = greenWhiteRed(50), textMatrix = textMatrix, setStdMargins = FALSE, cex.text = 0.4, zlim = c(-1,1), main = paste("Module-trait relationships"))
```

I really like how it looks

Gene significance and module membership:  
```{r}
table(moduleColors)
# modules Lightgreen (77 member genes) and grey60 (88 memeber genes) look the most interesting.
# Define variable sev containing the AverageSeverity column of datTrait
sev = as.data.frame(datTraits$AverageSeverity);
names(sev) = "severity"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership),nSamples))
names(geneModuleMembership) = paste("MM", modNames, sep="")
names(MMPvalue) = paste("p.MM", modNames, sep="")
geneTraitSignificance = as.data.frame(cor(datExpr, sev, use = "p"))
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))
names(geneTraitSignificance) = paste("GS.", names(sev), sep="")
names(GSPvalue) = paste("p.GS.", names(sev), sep="")
```

Looking deeper into module light green. Will look at grey60 later.

```{r}
module1 = "lightgreen"
column = match(module1, modNames);
moduleGenes = moduleColors==module1;
sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),abs(geneTraitSignificance[moduleGenes, 1]), xlab = paste("Module Membership in", module1, "module"), ylab = "Gene significance for tipburn severity", main = paste("Module membership vs. gene significance\n"), cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module1)
```

Hmm the possible candidate genes are not in module Lightgreen. Let's check module grey60:  

```{r}
module2 = "grey60"
column = match(module2, modNames);
moduleGenes = moduleColors==module2;
sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),abs(geneTraitSignificance[moduleGenes, 1]), xlab = paste("Module Membership in", module2, "module"), ylab = "Gene significance for tipburn severity", main = paste("Module membership vs. gene significance\n"), cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module2)
rownames(geneModuleMembership[moduleGenes, ])
```

Hmm not there either. Let's find out where they are:  
```{r}
int.genes = read.csv("Interesting_Genes.csv")
int = int.genes$Int.Gene
moduleInt = as.data.frame(moduleColors[is.na(match(rownames(geneModuleMembership),int)) == FALSE])
names(moduleInt) = "Module"
geneInt = geneModuleMembership[is.na(match(rownames(geneModuleMembership),int)) == FALSE,]
moduleInt$genes = rownames(geneInt)
moduleInt
```

Lsat_1_v5_gn_5_127021 (black module) has always been a gene of interest. (Plus he blue module in general is not that interesting...)

```{r}
module3 = "black"
column = match(module3, modNames);
moduleGenes = moduleColors==module3;
sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),abs(geneTraitSignificance[moduleGenes, 1]), xlab = paste("Module Membership in", module3, "module"), ylab = "Gene significance for tipburn severity", main = paste("Module membership vs. gene significance\n"), cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module3)
rownames(geneModuleMembership[moduleGenes, ])
```

Or purple (Lsat_1_v5_gn_5_129201)..?
```{r}
module4 = "purple"
column = match(module4, modNames);
moduleGenes = moduleColors==module4;
sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),abs(geneTraitSignificance[moduleGenes, 1]), xlab = paste("Module Membership in", module4, "module"), ylab = "Gene significance for tipburn severity", main = paste("Module membership vs. gene significance\n"), cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module4)
geneModuleMembership[moduleGenes, ]
```



